On First-order Proofs without syntax

### 2021-01-05 main.tex
Fix some typos

Let $\Gamma$ be a sequent in $\FOMLLm$. A \bfit{link} on $\Gamma$ is a pair of
leaves whose atoms are pre-dual. A \bfit{linking} on $\Gamma$ is a set of 
disjoint links such that each leaf of $\Gamma$ is in exacly one link. Similar
to the set of links in the linked fograph, a linking can be seen as a
unification problem, and a link is said \bfit{unifiable} if the corresponding
unification problem is solvable. \bfit{Dependencies} are defined as previously.

@@

A \bfit{switching graph} of a unification structure $\US(\lambda)$ is any
derivative of $\US(\lambda)$ obtained by keeping only one edge into each $\parr$
and $\forall$ and undirecting remaining edges. A linking is \bfit{correct} if it
is unifiable and all of the switching graphs of its associated unification
structure are acyclic.

@@

We also confound $\parr$ with $\vee$ and $\otimes$ with $\wedge$ as unification
nets and first-order nets (fonets) are defined in different contexts.

@@

\begin{definition} A \textit{switching path} of a  formula tree $\FT(A)$ is a
	path in $\FT(A)$ that does not go through both incoming edges of a $\parr$.
\end{definition}

@@

\begin{proposition}
If there exists an induced bimatching of the linked fograph $G = \Gr(A)$, then there exists a switching graph of the corresponding unification net which contains a cycle.
\begin{proof}
  Suppose that there exists a set $W$ inducing a bimatching in $G$.
  Then $(W, E_G)$ and $(W, L_G)$ are matchings.\\
  Let $E_W$ (resp. $L_W$) be the restriction of $E_G$ (resp. $L_G$) to $W$. \\
  If $E_W \cap L_W \neq \emptyset$, then there exist $u$ and $v$ such that $uv
  \in E_G$ and $uv \in L_G$. By Proposition \ref{prop2}, there exists a
  switching path of the formula tree of $A$. Together with the leap
  $uv$, this path induces a cycle in a switching graph of the
  corresponding unification structure. \\
  We can now suppose that $E_W$ and $L_W$ are disjoint. It is not difficult to
  see the existence of an alternating and elementary cycle in the bicoloured graph $(W, E_W \uplus L_W)$, i.e. a cycle of which the edges are alternately in $E_W$ and $L_W$ and containing no two equal vertices.
  By Proposition \ref{prop2}, this cycle induces a cycle in the unification
  structure. Now we want to construct a switching graph that contains this cycle.\\
  Consider a universal quantifier $\forall x$. If $\forall x \notin W$, then        
  we keep the incoming edge from its direct subformula and remove all the 
  dependencies. Otherwise, since $(W, L_G)$ is a matching, there exists a
  unique existential quantifier adjacent to $\forall x$ and we keep thus the
  corresponding edge in the unification structure. \\
  Now consider a $\parr$. We distinguish three cases:
  \begin{itemize}
    \item the cycle goes through none of the two branches (incoming edges) of
	    the $\parr$: we can choose an
	    arbitrary switching for this $\parr$
    \item the cycle goes through exactly one branch: we choose the corresponding
	    switching
    \item the cycle goes through both branches:
	    this means that there exist $v_L \in W$ (resp. $v_R$) in the left
		  (resp. right) branch, $u_L$, $u_R \in W$, such that
		  $u_Lv_L, u_Rv_R \in E_W$ and that the corresponding switching path from $u_L$ to $v_L$ (resp. from $u_R$ to $v_R$) goes through the left (resp. right) edge of $\parr$.

    \begin{center}	
    \begin{tikzpicture}
    \tikzstyle{every node}=[circle, minimum size=12pt, inner sep=2pt]
    \draw (0, 0) node (par) {$\parr$};
    \draw (1, 1.7) node (dots1) {$\vdots$};
    \draw (-1, 1.7) node (dots2) {$\vdots$};
    \draw (1, 3.4) node (vr) {$v_R$};
    \draw (-1, 3.4) node (vl) {$v_L$};
    \draw (1.5, -1) node (dots3) {$\vdots$};
    \draw (-1.5, -1) node (dots4) {$\vdots$};
    \draw (2.9, -1) node (ul) {$u_L$};
    \draw (-2.9, -1) node (ur) {$u_R$};
    \draw (par) -- (dots1) [color=blue];
    \draw (par) -- (dots2) [color=red];
    \draw (dots1) -- (vr) [color=blue]; 
    \draw (dots2) -- (vl) [color=red];
    \draw (dots4) -- (ur) [color=blue]; 
    \draw (dots3) -- (ul) [color=red];
    \draw (dots4) -- (par) [color=blue]; 
    \draw (dots3) -- (par) [color=red];
    \end{tikzpicture}

    {\bf Figure 7.} A schema showing that the two branches of the same $\parr$ cannot be used in the cycle at the same time. 
    \end{center}

@@

    It is clear that $u_L$ (resp. $u_R$) is not in the branches of the
	 $\parr$. Otherwise, there will be no switching path from $u_L$ to $v_L$

    By Proposition \ref{prop2}, we know that $u_L$ and $u_R$ are not
	universal quantifiers which are ancestors the $\parr$
	and that there exist one switching path from $u_L$ to $v_L$ and one
	from $u_R$ to $v_R$. In particular, there exist one switching path from
	$u_L$ to the $\parr$ and one from the $\parr$ to $v_R$, and by
	concatenating the two, we obtain a switching path from $u_L$ to $v_R$. 
	By Proposition \ref{prop2}, $u_L$ and $v_R$ are thus adjacent in $(W,
	E_G)$, which is impossible since $(W, E_W)$ is a matching.
  \end{itemize}
  Notice that the switching paths here are in the underlying formula tree. We
  have to verify that they are compatible with the choices of switching made for
  universal quantifiers. That is, if $uv \in E_W$, then for all the
  universal quantifiers $\forall x$ on the switching path (in the formula
	tree), we have chosen in the first part of the proof to keep the only
	edge from the child of $\forall x$ to itself. In fact, if there exists
	a universal quantifier $w \in W$ on the switching path $u \rightarrow
	v$, then one of $u$ and $v$ is not a descendant of $w$. Moreover, if
	$u$ (resp. $v$) is a universal quantifier, then $w$ is not in its scope.
	By Proposition \ref{prop2}, $\{wu, wv\} \cap E_W \neq \emptyset$, which
	is impossible since $(W, E_W)$ is a matching.
  We have thus constructed a switching graph containing this cycle.
\end{proof}
\end{proposition}

@@

\begin{enumerate}
  \item {\bf Encode dependencies as fresh links.} For each dependency $\exists x \rightarrow \forall y$, with corresponding subformulas $\exists x A$ and $\forall y B$, we add a fresh link as follows. Let $P$ be a fresh (nullary) predicate symbol. Replace $\exists x A$ with $P \otimes \exists x A$ and $\forall y B$ with $\overline{P} \parr \forall y B$, and add an axiom link between $P$ and $\overline{P}$.
  \item {\bf Erase quantifiers.} After step 1, erase all the quantifiers. (We no longer need their leaps since they are encoded as links in step 1.
  \item {\bf Simplify atoms.} After step 2, replace every predicate $Pt_1 \cdots t_n$ with a nullary predicate symbol $P$.
\end{enumerate}
\end{definition}

\begin{center}
\begin{tikzpicture}
  \tikzstyle{every node}=[circle, minimum size=20pt, inner sep=2pt]
  \draw (0, 0) node (par1) {$\parr$};
  \draw (1.8, 1.2) node (z) {\color{blue} $\forall z$};
  \draw (-1.8, 1.2) node (par2) {$\parr$};
    \draw (-2.8, 2.4) node (x) {\color{red} $\exists x$};
    \draw (-0.8, 2.4) node (y) {$\exists y$};
    \draw (-2.8, 3.6) node (px) {$\overline{p}x$};
    \draw (-0.8, 3.6) node (qy) {$\overline{q}y$};
    \draw (1.8, 2.4) node (tens) {$\otimes$};
    \draw (0.8, 3.6) node (pz) {$pz$};
    \draw (2.8, 3.6) node (qfz) {$qfz$};
    \draw (par1) -- (par2) [<-];
    \draw (par1) -- (z) [<-];
    \draw (par2) -- (x) [<-]; 
    \draw (par2) -- (y) [<-];
    \draw (x) -- (px) [<-]; 
    \draw (y) -- (qy) [<-];
    \draw (z) -- (tens) [<-]; 
    \draw (tens) -- (pz) [<-];
    \draw (tens) -- (qfz) [<-];
    \draw (y) -- (z) [dotted, ->];
    \draw (x) -- (z) [dotted, ->];
    \draw[color=blue, dashed] (px) to [bend left] (pz);
    \draw[color=red, dashed] (qy) to [bend left] (qfz);
\end{tikzpicture}
\vspace{0.5cm}

\begin{tikzpicture}
  \tikzstyle{every node}=[circle, minimum size=20pt, inner sep=2pt]
  \draw (0, 0) node (par1) {$\parr$};
  \draw (1, 0.9) node (par3) {\color{blue} $\parr$};
  \draw (-1.8, 1.2) node (par2) {$\parr$};
  \draw (-2.6, 2.4) node (tens1) {\color{red} $\otimes$};
  \draw (-1, 2.4) node (tens2) {$\otimes$};
  \draw (-2.2, 3.6) node (S) {\color{red} $S$};
  \draw (-0.6, 3.6) node (R) {$R$};
  \draw (-3, 3.6) node (nP) {$\overline{P}$};
  \draw (-1.4, 3.6) node (nQ) {$\overline{Q}$};
  \draw (0.4, 3.6) node (nS) {\color{blue} $\overline{S}$};
  \draw (1.6, 1.8) node (par4) {$\parr$};
  \draw (2.2, 2.7) node (tens3) {$\otimes$};
  \draw (2, 3.6) node (P) {$P$};
  \draw (2.8, 3.6) node (Q) {$Q$};
  \draw (1.2, 3.6) node (nR) {$\overline{R}$};
  \draw (par2) -- (par1) [->];
  \draw (par3) -- (par1) [->];
  \draw (tens1) -- (par2) [->];
  \draw (tens2) -- (par2) [->];
  \draw (nP) -- (tens1) [->];
  \draw (S) -- (tens1) [->, color=red];
  \draw (nQ) -- (tens2) [->];
  \draw (R) -- (tens2) [->];
  \draw (nS) -- (par3) [->, color=blue];
  \draw (nR) -- (par4) [->];
  \draw (tens3) -- (par4) [->];
  \draw (P) -- (tens3) [->];
  \draw (Q) -- (tens3) [->];
  \draw (par4) -- (par3) [->];
  \draw (nP) to [bend left] (P);
  \draw (nQ) to [bend left] (Q);
  \draw (S) to [bend left] (nS);
  \draw (R) to [bend left] (nR);
\end{tikzpicture}

>>>>

Let $\Gamma$ be a sequent in $\FOMLLm$. A \bfit{link} on $\Gamma$ is a pair of
leaves whose atoms are pre-dual. A \bfit{linking} on $\Gamma$ is a set of 
disjoint links such that each leaf of $\Gamma$ is in exactly one link. Similar
to the set of links in the linked fograph, a linking can be seen as a
unification problem, and a link is said \bfit{unifiable} if the corresponding
unification problem is solvable. \bfit{Dependencies} are defined as previously.

@@

A \bfit{switching graph} of a unification structure $\US(\lambda)$ is any
derivative of $\US(\lambda)$ obtained by keeping only one edge into each $\vee$
and $\forall$ and undirecting remaining edges. A linking is \bfit{correct} if it
is unifiable and all of the switching graphs of its associated unification
structure are acyclic.

@@

@@

\begin{definition} A \textit{switching path} of a  formula tree $\FT(A)$ is a
	path in $\FT(A)$ that does not go through both incoming edges of a $\vee$.
\end{definition}

@@

\begin{proposition}
If there exists an induced bimatching of the linked fograph $G = \Gr(A)$, then there exists a switching graph of the corresponding unification net which contains a cycle.
\begin{proof}
  Suppose that there exists a set $W$ inducing a bimatching in $G$.
  Then $(W, E_G)$ and $(W, L_G)$ are matchings.\\
  Let $E_W$ (resp. $L_W$) be the restriction of $E_G$ (resp. $L_G$) to $W$. \\
  If $E_W \cap L_W \neq \emptyset$, then there exist $u$ and $v$ such that $uv
  \in E_G$ and $uv \in L_G$. By Proposition \ref{prop2}, there exists a
  switching path of the formula tree of $A$. Together with the leap
  $uv$, this path induces a cycle in a switching graph of the
  corresponding unification structure. \\
  We can now suppose that $E_W$ and $L_W$ are disjoint. It is not difficult to
  see the existence of an alternating and elementary cycle in the bicoloured graph $(W, E_W \uplus L_W)$, i.e. a cycle of which the edges are alternately in $E_W$ and $L_W$ and containing no two equal vertices.
  By Proposition \ref{prop2}, this cycle induces a cycle in the unification
  structure. Now we want to construct a switching graph that contains this cycle.\\
  Consider a universal quantifier $\forall x$. If $\forall x \notin W$, then        
  we keep the incoming edge from its direct subformula and remove all the 
  dependencies. Otherwise, since $(W, L_G)$ is a matching, there exists a
  unique existential quantifier adjacent to $\forall x$ and we keep thus the
  corresponding edge in the unification structure. \\
  Now consider a $\vee$. We distinguish three cases:
  \begin{itemize}
    \item the cycle goes through none of the two branches (incoming edges) of
	    the $\vee$: we can choose an
	    arbitrary switching for this $\vee$
    \item the cycle goes through exactly one branch: we choose the corresponding
	    switching
    \item the cycle goes through both branches:
	    this means that there exist $v_L \in W$ (resp. $v_R$) in the left
		  (resp. right) branch, $u_L$, $u_R \in W$, such that
		  $u_Lv_L, u_Rv_R \in E_W$ and that the corresponding switching
		  path from $u_L$ to $v_L$ (resp. from $u_R$ to $v_R$) goes
		  through the left (resp. right) edge of $\vee$.

    \begin{center}	
    \begin{tikzpicture}
    \tikzstyle{every node}=[circle, minimum size=12pt, inner sep=2pt]
    \draw (0, 0) node (par) {$\vee$};
    \draw (1, 1.7) node (dots1) {$\vdots$};
    \draw (-1, 1.7) node (dots2) {$\vdots$};
    \draw (1, 3.4) node (vr) {$v_R$};
    \draw (-1, 3.4) node (vl) {$v_L$};
    \draw (1.5, -1) node (dots3) {$\vdots$};
    \draw (-1.5, -1) node (dots4) {$\vdots$};
    \draw (2.9, -1) node (ul) {$u_L$};
    \draw (-2.9, -1) node (ur) {$u_R$};
    \draw (par) -- (dots1) [color=blue];
    \draw (par) -- (dots2) [color=red];
    \draw (dots1) -- (vr) [color=blue]; 
    \draw (dots2) -- (vl) [color=red];
    \draw (dots4) -- (ur) [color=blue]; 
    \draw (dots3) -- (ul) [color=red];
    \draw (dots4) -- (par) [color=blue]; 
    \draw (dots3) -- (par) [color=red];
    \end{tikzpicture}

    {\bf Figure 7.} A schema showing that the two branches of the same $\vee$ cannot be used in the cycle at the same time. 
    \end{center}

@@

    It is clear that $u_L$ (resp. $u_R$) is not in the branches of the
	 $\vee$. Otherwise, there will be no switching path from $u_L$ to $v_L$

    By Proposition \ref{prop2}, we know that $u_L$ and $u_R$ are not
	universal quantifiers which are ancestors the $\vee$
	and that there exist one switching path from $u_L$ to $v_L$ and one
	from $u_R$ to $v_R$. In particular, there exist one switching path from
	$u_L$ to the $\vee$ and one from the $\vee$ to $v_R$, and by
	concatenating the two, we obtain a switching path from $u_L$ to $v_R$. 
	By Proposition \ref{prop2}, $u_L$ and $v_R$ are thus adjacent in $(W,
	E_G)$, which is impossible since $(W, E_W)$ is a matching.
  \end{itemize}
  Notice that the switching paths here are in the underlying formula tree. We
  have to verify that they are compatible with the choices of switching made for
  universal quantifiers. That is, if $uv \in E_W$, then for all the
  universal quantifiers $\forall x$ on the switching path (in the formula
	tree), we have chosen in the first part of the proof to keep the only
	edge from the child of $\forall x$ to itself. In fact, if there exists
	a universal quantifier $w \in W$ on the switching path $u \rightarrow
	v$, then one of $u$ and $v$ is not a descendant of $w$. Moreover, if
	$u$ (resp. $v$) is a universal quantifier, then $w$ is not in its scope.
	By Proposition \ref{prop2}, $\{wu, wv\} \cap E_W \neq \emptyset$, which
	is impossible since $(W, E_W)$ is a matching.
  We have thus constructed a switching graph containing this cycle.
\end{proof}
\end{proposition}

@@

\begin{enumerate}
  \item {\bf Encode dependencies as fresh links.} For each dependency $\exists x
	  \rightarrow \forall y$, with corresponding subformulas $\exists x A$
		and $\forall y B$, we add a fresh link as follows. Let $P$ be a
		fresh (nullary) predicate symbol. Replace $\exists x A$ with $P
		\wedge \exists x A$ and $\forall y B$ with $\overline{P} \vee \forall y B$, and add an axiom link between $P$ and $\overline{P}$.
  \item {\bf Erase quantifiers.} After step 1, erase all the quantifiers. (We no longer need their leaps since they are encoded as links in step 1.
  \item {\bf Simplify atoms.} After step 2, replace every predicate $Pt_1 \cdots t_n$ with a nullary predicate symbol $P$.
\end{enumerate}
\end{definition}

\begin{center}
\begin{tikzpicture}
  \tikzstyle{every node}=[circle, minimum size=20pt, inner sep=2pt]
  \draw (0, 0) node (par1) {$\vee$};
  \draw (1.8, 1.2) node (z) {\color{blue} $\forall z$};
  \draw (-1.8, 1.2) node (par2) {$\vee$};
    \draw (-2.8, 2.4) node (x) {\color{red} $\exists x$};
    \draw (-0.8, 2.4) node (y) {$\exists y$};
    \draw (-2.8, 3.6) node (px) {$\overline{p}x$};
    \draw (-0.8, 3.6) node (qy) {$\overline{q}y$};
    \draw (1.8, 2.4) node (tens) {$\wedge$};
    \draw (0.8, 3.6) node (pz) {$pz$};
    \draw (2.8, 3.6) node (qfz) {$qfz$};
    \draw (par1) -- (par2) [<-];
    \draw (par1) -- (z) [<-];
    \draw (par2) -- (x) [<-]; 
    \draw (par2) -- (y) [<-];
    \draw (x) -- (px) [<-]; 
    \draw (y) -- (qy) [<-];
    \draw (z) -- (tens) [<-]; 
    \draw (tens) -- (pz) [<-];
    \draw (tens) -- (qfz) [<-];
    \draw (y) -- (z) [dotted, ->];
    \draw (x) -- (z) [dotted, ->];
    \draw[color=blue, dashed] (px) to [bend left] (pz);
    \draw[color=red, dashed] (qy) to [bend left] (qfz);
\end{tikzpicture}
\vspace{0.5cm}

\begin{tikzpicture}
  \tikzstyle{every node}=[circle, minimum size=20pt, inner sep=2pt]
  \draw (0, 0) node (par1) {$\vee$};
  \draw (1, 0.9) node (par3) {\color{blue} $\vee$};
  \draw (-1.8, 1.2) node (par2) {$\vee$};
  \draw (-2.6, 2.4) node (tens1) {\color{red} $\wedge$};
  \draw (-1, 2.4) node (tens2) {$\wedge$};
  \draw (-2.2, 3.6) node (S) {\color{red} $S$};
  \draw (-0.6, 3.6) node (R) {$R$};
  \draw (-3, 3.6) node (nP) {$\overline{P}$};
  \draw (-1.4, 3.6) node (nQ) {$\overline{Q}$};
  \draw (0.4, 3.6) node (nS) {\color{blue} $\overline{S}$};
  \draw (1.6, 1.8) node (par4) {$\vee$};
  \draw (2.2, 2.7) node (tens3) {$\wedge$};
  \draw (2, 3.6) node (P) {$P$};
  \draw (2.8, 3.6) node (Q) {$Q$};
  \draw (1.2, 3.6) node (nR) {$\overline{R}$};
  \draw (par2) -- (par1) [->];
  \draw (par3) -- (par1) [->];
  \draw (tens1) -- (par2) [->];
  \draw (tens2) -- (par2) [->];
  \draw (nP) -- (tens1) [->];
  \draw (S) -- (tens1) [->, color=red];
  \draw (nQ) -- (tens2) [->];
  \draw (R) -- (tens2) [->];
  \draw (nS) -- (par3) [->, color=blue];
  \draw (nR) -- (par4) [->];
  \draw (tens3) -- (par4) [->];
  \draw (P) -- (tens3) [->];
  \draw (Q) -- (tens3) [->];
  \draw (par4) -- (par3) [->];
  \draw (nP) to [bend left] (P);
  \draw (nQ) to [bend left] (Q);
  \draw (S) to [bend left] (nS);
  \draw (R) to [bend left] (nR);
\end{tikzpicture}
